---
title: 'P&S-2022: Lab assignment 2'
author: "Name1, Name2, Viktoria Toporovska"
output:
  html_document:
    df_print: paged
---

### Task 3.

#### In this task, we use the Central Limit Theorem approximation for continuous random variables.

#### One of the devices to measure radioactivity level at a given location is the Geiger counter. When the radioactive level is almost constant, the time between two consecutive clicks of the Geiger counter is an exponentially distributed random variable with parameter $\nu_1 = \mathtt{team\,id\,number} + 10$. Denote by $X_k$ the random time between the $(k-1)^{\mathrm{st}}$ and $k^{\mathrm{th}}$ click of the counter.

1.  Show that the distribution of the sample means of $X_1, X_2,\dots,X_n$ gets very close to a normal one (which one?) as $n$ becomes large. To this end,
    -   simulate the realizations $x_1,x_2,\dots,x_n$ of the \textbf{r.v.} $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;\
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and then the \emph{empirical cumulative distribution} function $F_{\mathbf{s}}$ of $\mathbf{s}$;\
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} $F_{\mathbf{s}}$ of and plot both \textbf{c.d.f.}'s on one graph to visualize their proximity;\
    -   calculate the maximal difference between the two \textbf{c.d.f.}'s;\
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.

#### First, generate samples an sample means:

```{r}
# id = 3
nu1 <- 13  # id + 10
K <- 1e3 # times to get the sample
n_values <- c(5, 10, 50)  # Sample sizes

n = 50 #test or start size

sample_means <- matrix(rexp(n * K, rate = nu1), nrow = n)
```

#### Next, calculate the parameters of the standard normal approximation

```{r}
mu <- 1/nu1 #we are looking for limits
sigma <-  mu / sqrt(n) 
print(mu)
print (sigma)
```

#### We can now plot ecdf and cdf

```{r}
x <- seq(-1,1.200,by=0.001)# set the dimension of the x axis
for (n in n_values) { #passability of the cycle on all elements
  mu <- 1 / nu1 #we are looking for everyone
  sigma <- mu / sqrt(n)
  
  xlims <- c(mu-3*sigma,mu+3*sigma) 
  

  sample_means <- colMeans(matrix(rexp(n*K, rate = nu1), nrow=n))
  Fs <- ecdf(sample_means)
  
  title <- paste("comparing cdf and ecdf for", toString(n))
  plot(Fs, 
       xlim = xlims, 
       col = "blue",
       lwd = 2,
       main = title)
  
  curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)
  
  print(max(abs(ecdf(sample_means)(x)-pnorm(x,mean = mu, sd = sigma))))
}
```

Normal one refers to the standard normal distribution with $\mu = 0$ and $\sigma = 1$. With the help of visualization, you can observe how much the two graphs converge as n increases.

1.  The place can be considered safe when the number of clicks in one minute does not exceed $100$. It is known that the parameter $\nu$ of the resulting exponential distribution is proportional to the number $N$ of the radioactive samples, i.e., $\nu = \nu_1*N$, where $\nu_1$ is the parameter for one sample. Determine the maximal number of radioactive samples that can be stored in that place so that, with probability $0.95$, the place is identified as safe. To do this,
    -   express the event of interest in terms of the \textbf{r.v.} $S:= X_1 + \cdots + X_{100}$;
    -   obtain the theoretical bounds on $N$ using the Markov inequality, Chernoff bound and Central Limit Theorem and compare the results;
    -   with the predicted $N$ and thus $\nu$, simulate the realization $x_1,x_2,\dots,x_{100}$ of the $X_i$ and of the sum $S = X_1 + \cdots + X_{100}$;
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of total times until the $100^{\mathrm{th}}$ click;
    -   estimate the probability that the location is identified as safe and compare to the desired level $0.95$

$S = X_1 + ... + X_{100}$

$P(S>=60)$

**Markov** $$
P(S \ge 1) \le \frac {\mu_S}{1} = \frac {100} {13*N}
$$ $$
0.95 \le \frac {100} {13*N}
$$ $$N=8
$$

**Central Limit theorem:**

$$
P(\frac {S - 100 * \mu} {\sigma * \sqrt n} \le t) \rightarrow Ф(t)
$$

$$
\mu_S = \frac {100} {13*N} = \sigma_S
$$ $$
P(S\ge1) = P(\frac {S - 100 * \frac 1 {13 * N}}{10 * \frac 1 {13 * N}} \le \frac {1 - 100 * \frac 1 {13 * N}}{10 * \frac 1 {13 * N}}) = Ф(\frac {13 * N- 100} {10})
$$ $$
Ф(\frac {13 * N- 100} {10}) \ge 0.95
$$

```{r}
threshold <- 60  # Threshold for safety (total time <= 60 seconds)
desired_prob <- 0.95  # Desired probability of safety

# Step 1: Express the event of interest
# Calculate the probability that the total time taken to accumulate 100 clicks is more than 60 seconds
S_event <- rexp(100, rate = nu1)
S <- sum(S_event)

# Step 2: Theoretical Bounds
# Calculate theoretical bounds on N using Markov, Chernoff, and CLT
# Markov Bound
markov_bound <- 100 / (desired_prob * nu1)  # Upper bound based on Markov inequality

# Chernoff Bound
t <- log(1 / (desired_prob ^ (1/100)))  # Chernoff bound parameter
chernoff_bound <- (t * (exp(1) * desired_prob) ^ (1/100)) / (13 * ((exp(1) * desired_prob) ^ (1/100) - 1))

# CLT Bound
clt_bound <- 100 / (13 * markov_bound)  # Upper bound based on CLT

# Output bounds
cat("Markov Bound:", markov_bound, "\n")
cat("Chernoff Bound:", chernoff_bound, "\n")
cat("CLT Bound:", clt_bound, "\n")

```

```{r}

```
