```{r}
id <- 3
set.seed(id)
```

**Do not forget to include several sentences summarizing your work and the conclusions you have made!**

### Task 2.

#### In this task, we discuss a real-life process that is well modelled by a Poisson distribution. As you remember, a Poisson random variable describes occurrences of rare events, i.e., counts the number of successes in a large number of independent random experiments. One of the typical examples is the **radioactive decay** process.

#### Consider a sample of radioactive element of mass $m$, which has a big *half-life period* $T$; it is vitally important to know the probability that during a one second period, the number of nuclei decays will not exceed some critical level $k$. This probability can easily be estimated using the fact that, given the *activity* ${\lambda}$ of the element (i.e., the probability that exactly one nucleus decays in one second) and the number $N$ of atoms in the sample, the random number of decays within a second is well modelled by Poisson distribution with parameter $\mu:=N\lambda$. Next, for the sample of mass $m$, the number of atoms is $N = \frac{m}{M} N_A$, where $N_A = 6 \times 10^{23}$ is the Avogadro constant, and $M$ is the molar (atomic) mass of the element. The activity of the element, $\lambda$, is $\log(2)/T$, where $T$ is measured in seconds.

#### Assume that a medical laboratory receives $n$ samples of radioactive element ${{}^{137}}\mathtt{Cs}$ (used in radiotherapy) with half-life period $T = 30.1$ years and mass $m = \mathtt{team\, id \,number} \times 10^{-6}$ g each. Denote by $X_1,X_2,\dots,X_n$ the **i.i.d. r.v.**'s counting the number of decays in sample $i$ in one second.

1.  Specify the parameter of the Poisson distribution of $X_i$ (you'll need the atomic mass of *Cesium-137*)\

```{r}

m <- id/10^6            
T <- 30.1*365*24*60*60  
lambda <- log(2)/T  
N <- m * 6 * 10^(23) / (137)
mu <- N * lambda  
print(mu)
```

2.  Show that the distribution of the sample means of $X_1,\dots,X_n$ gets very close to a normal one as $n$ becomes large and identify that normal distribution. To this end,
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and form the empirical cumulative distribution function $\hat F_{\mathbf{s}}$ of $\mathbf{s}$;
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} $F$ of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} $\hat F_{\mathbf{s}}$ and plot both **c.d.f.**'s on one graph to visualize their proximity (use the proper scales!);
    -   calculate the maximal difference between the two \textbf{c.d.f.}'s;
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.\

```{r}
sample_means <- colMeans(matrix(rpois(n*K, lambda = mu), nrow=n))
mu <- mean(sample_means)
sigma <- sd(sample_means)
```

For n = 5:

```{r}
xlims <- c(mu - 3 * sigma, mu + 3 * sigma)
Fs <- ecdf(sample_means)

x <- seq(mu - 3 * sigma, mu + 3 * sigma, by = 1)
print(max(abs(Fs(x) - pnorm(x, mean = mu, sd = sigma))))

plot(Fs,
     xlim = xlims,
     ylim = c(0, 1),
     col = "green",
     lwd = 1,
     main = "Comparison of ecdf and cdf for n = 5",
     xlab = "Value",
     ylab = "Probability")

curve(pnorm(x, mean = mu, sd = sigma),
       col = "blue",
       lwd = 2,
       add = TRUE)

legend("topleft",
       c("ecdf", "cdf"),
       col = c("green", "blue"),
       lwd = c(1, 2))

```

For n = 10:

```{r}
n <- 10
K <- 1000
sample_means <- colMeans(matrix(rpois(n * K, lambda = mu), nrow = n))
mu <- mean(sample_means)
sigma <- sd(sample_means)
xlims <- c(mu - 3 * sigma, mu + 3 * sigma)
Fs <- ecdf(sample_means)

x <- seq(mu - 3 * sigma, mu + 3 * sigma, by = 1)

se <- sd(sample_means) / sqrt(n)
lower_bound <- Fs(x) - se
upper_bound <- Fs(x) + se

plot(Fs,
     xlim = xlims,
     ylim = c(0, 1),
     col = "blue",
     lwd = 2,
     main = "Comparison of ecdf and cdf for n = 10")
curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)

```

For n = 50:

```{r}
n <- 500
sample_means <- colMeans(matrix(rpois(n * K, lambda = mu), nrow = n))
mu <- mean(sample_means)
sigma <- sd(sample_means)
xlims <- c(mu - 3 * sigma, mu + 3 * sigma)
Fs <- ecdf(sample_means)

x <- seq(mu - 3 * sigma, mu + 3 * sigma, by = 1)

# Змініть кольори і стиль ліній для графіків
plot(Fs, 
     xlim = xlims, 
     ylim = c(0, 1),
     col = "green", 
     lwd = 2,       
     main = "Comparison of ecdf and cdf for n = 50")
curve(pnorm(x, mean = mu, sd = sigma), col = "blue", lwd = 2, add = TRUE)  # Синій колір
```

3.  Calculate the largest possible value of $n$, for which the total number of decays in one second is less than $8 \times 10^8$ with probability at least $0.95$. To this end,
    -   obtain the theoretical bound on $n$ using Markov inequality, Chernoff bound and Central Limit Theorem, and compare the results;\
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sum $s=x_1 + \cdots +x_n$;
    -   repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of sums;
    -   calculate the number of elements of the sample which are less than critical value ($8 \times 10^8$) and calculate the empirical probability; comment whether it is close to the desired level $0.95$

### Markov's inequality:

$$
P(X >= 8 * 10^8) <= \frac{\mu n}{(8 * 10^8)} <= 0.05
$$ $$
\mu = 9594101, n <= \frac{0.05 * 8 * 10^8}\mu = \frac{0.05 * 8 * 10^8}{9594101} = 4.16922857
$$

### Chernoff inequality:

$$
P(X>=t)<= \frac {e^{-\lambda}(e\lambda)^t}{t^t}
$$

$$
P(X>=8*10^8) <= \frac {e^{9594101n}*(9594101en)^{8*10^8}}{(8*10^8)^{8*10^8}} <= 0.05
$$ $$
e^{9594101n}*n^{8*10^8} <= \frac {0.05 * (8*10^8)^{8*10^8}}{e^{8*10^8*9594101}}
$$

### CLT

$$
P(\frac{X_1 + X_2+...+X_n  - \mu n}{\sigma \sqrt n} <= t) \longrightarrow \Phi(t)
$$

$$
t = \frac {8 * 10^8 - 9594101n}{\sqrt {9594101n}}
$$

$$
\Phi(t) >= 0.95
$$

$$
t >= \Phi^{-1}(0.95) = 1 - \Phi(0.95) = 0.1790561
$$

$$
\frac {8 * 10^8 - 9594101n}{\sqrt {9594101n}} >= 0.1790561
$$

$$
8 * 10^8 >= \sqrt {9594101n} + 9594101n
$$

$$
n <= 83.38
$$
